\textbf{COMP0086 Summative Assignment}

### 2022-11-14 submission.tex
insert missing negative in log bits

$$\log_2(P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta))$$

@@

$$rate = 1 - \frac{\log_2(P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta))}{N \cdot D}$$

@@

$$\log_2(P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta)) + M \cdot K \cdot D + M \cdot K$$

Where $M$ is the cost of storing a single float value, in our case we used $float64$ so 64 bits. The first term is the log-likelihood as expressed in part (f), the second term is the cost of storing $\mathbf{P}$, and the last term is the cost of storing $\pi$. The latter two terms scale with the value of k. This means that as $k$ increases, our compression rate deteriorates. Looking at the total compression ratio $\frac{log_2(P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta)) + M \cdot K \cdot D + M \cdot K}{N\cdot D}$ in a table:

@@

We can see that the ratio is greater than one, meaning that this is actually worse than the naive encoding. This is due to the high cost of storing each value in the model parameters being 64 bits. However, because this remains constant with respect to our data set size, as $N$ increases these ratios will approach $\frac{\log_2(P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta)}{N\cdot D}$ and we'll recover our compression rates from part (f). By increasing $k$, we see that the ratio increases, and thus our compression rate is worse. This makes sense because we are essentially slowly storing the data into the model. In the extreme example of $k=N$ we can have the parameters for each mixture model as an image in the data. Although in this case, we could store our parameters as binary values instead of 64 bit floats. Our mixture component is uniform because each mixture is equally likely (all images are equally likely) so there is no need to store any values for $\pi$. Thus, we would recover the cost of naive encoding our data. Taking into account the model parameters further verifies that the cost of encoding with this approach is much higher than the cost of gzip.

>>>>

$$-\log_2(P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta))$$

@@

$$rate = 1 - \frac{-\log_2(P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta))}{N \cdot D}$$

@@

$$-\log_2(P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta)) + M \cdot K \cdot D + M \cdot K$$

Where $M$ is the cost of storing a single float value, in our case we used $float64$ so 64 bits. The first term is the log-likelihood as expressed in part (f), the second term is the cost of storing $\mathbf{P}$, and the last term is the cost of storing $\pi$. The latter two terms scale with the value of k. This means that as $k$ increases, our compression rate deteriorates. Looking at the total compression ratio $\frac{-log_2(P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta)) + M \cdot K \cdot D + M \cdot K}{N\cdot D}$ in a table:

@@

We can see that the ratio is greater than one, meaning that this is actually worse than the naive encoding. This is due to the high cost of storing each value in the model parameters being 64 bits. However, because this remains constant with respect to our data set size, as $N$ increases these ratios will approach $\frac{-\log_2(P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta)}{N\cdot D}$ and we'll recover our compression rates from part (f). By increasing $k$, we see that the ratio increases, and thus our compression rate is worse. This makes sense because we are essentially slowly storing the data into the model. In the extreme example of $k=N$ we can have the parameters for each mixture model as an image in the data. Although in this case, we could store our parameters as binary values instead of 64 bit floats. Our mixture component is uniform because each mixture is equally likely (all images are equally likely) so there is no need to store any values for $\pi$. Thus, we would recover the cost of naive encoding our data. Taking into account the model parameters further verifies that the cost of encoding with this approach is much higher than the cost of gzip.
